\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bettelini}
\usepackage{stellar}
\usepackage{definitions}

\begin{document}

\id{english-frankenstein-chapter-17-18}
\genpage

\subsection{Exercises}

\begin{snippetexercise}{frankenstein-ex-33}
    {What persuades Frankenstein to create a companion for the creature?}
    The monster blackmails Frankenstein by saying that if he didn't create a female-monster,
    then he would destroy the human beings 
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-34}
    {How does Victor react when his father brings up the subject of his expected marriage with
    Elizabeth?}
    Initially, he says that despite the fact that he thinks very high of her, he doesn't want
    to get married in a short time because he's afraid that the monster could 
    harm her and their family as well.
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-35}
    {What is Victor worried about before he leaves for England?}
    He's worried because he doesn't want to leave his family behind
    with the risk of them getting killed by the monster.
    He's also worried because he's going with his friend Clerval and he could 
    slow down his work.
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-36}
    {What do the last lines of Chapter 18 imply about the potential fate awaiting Frankenstein's friend
    Clerval?}
    Frankenstein anticipates that his friend Clerval is dead, but he doesn't tell us why,
    how or when. he only talks about his feelings. 
\end{snippetexercise}

\end{document}
