\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stellar}
\usepackage{bettelini}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Stellar},
    pdfpagemode=FullScreen,
}

\begin{document}

\id{english-aviewfromthebridge-act-two-part-1}
\genpage

\section{Exercises}

\begin{snippetexercise}{aviewfromthebridge-ex-23}
    {At the beginning of Act two Alfieri briefly sets up the scene. What does he narrate that is relevant
    for the continuation of the story?}
    TODO
\end{snippetexercise}

\begin{snippetexercise}{aviewfromthebridge-ex-24}
    {What does Catherine ask Rodolpho that makes her look worried?}
    TODO

    \begin{enumerate}
        \item \textbf{Are his reasons for not wanting to go back to Italy?}
        \item \textbf{Why does Catherine ask Rodolpho about going back to Italy? How does he react to her
        questions?}
    \end{enumerate}
\end{snippetexercise}

\begin{snippetexercise}{aviewfromthebridge-ex-25}
    {Read the following excerpt from Act two. What does it reveal about Catherin?}
    \hspace{0.1\textwidth}
    \begin{minipage}[r]{0.8\textwidth}
        \itshape
        If I was a wife I would make a man happy instead of goin' at him all the time. I can tell a
        block away when he's blue in his mind and just wants to talk to somebody quiet and
        nice… I can tell when he's hungry or wants a beer before he even says anything. I know
        when his feet hurt him, I mean I know him and now I'm supposed to turn around and
        make a stranger out of him? I don't know why I have to do that, I mean.
    \end{minipage}
    \\\\
    TODO
\end{snippetexercise}

\begin{snippetexercise}{aviewfromthebridge-ex-26}
    {What is Rodolpho trying to explain to her using the metaphor of the bird?}
    TODO
\end{snippetexercise}

\begin{snippetexercise}{aviewfromthebridge-ex-27}
    {Towards the end of the scene Eddie arrives home drunk while Catherine and Rodolpho were being
    intimate. How does he react to stumbling upon this?}
    TODO
\end{snippetexercise}

\begin{snippetexercise}{aviewfromthebridge-ex-28}
    {Why does Eddie kiss both Catherine and Rodolpho?}
    TODO
\end{snippetexercise}

\end{document}