\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bettelini}
\usepackage{stellar}
\usepackage{definitions}
\usepackage{makecell}

\begin{document}

\id{english-atonement-chapter8-10}
\genpage

\section{Summary}

\section{Adjectives}

\begin{snippet}{atonement-chapter8-10-adjectives}
\begin{minipage}[l]{0.05\textwidth}
    \circled{1}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Adjective:} \textit{awkward}
    \\
    \textbf{Who / What?:} This is referring to Cecilia and her relationship with Robbie.
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{2}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Adjective:} \textit{restless}
    \\
    \textbf{Who / What?:} Cecilia is trying off the dress and she keeps moving around.
    She cannot stay still at home.
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{3}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Adjective:} \textit{angry}
    \\
    \textbf{Who / What?:} This is also referring to Cecilia, which was angry because
    she did not want Robbie to accept the invitation from Paul.
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{4}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Adjective:} \textit{clumsy and inconsiderate}
    \\
    \textbf{Who / What?:} This is Robbie describing his own behavior
    at the fountain episode.
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{5}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Adjective:} \textit{horrified}
    \\
    \textbf{Who / What?:} Robbie feels horrified when he realized he had given
    the wrong letter to Briony.
\end{minipage}
\end{snippet}

\section{Exercises}

\begin{snippetexercise}{atonement-ex-5}%
{What is the relevance of the scene where Cecilia reflects on what outfit to wear for dinner?}
Cecilia doesn't want to show off to much, so she overthinks about the kind of dress,
but she also wants to show that she didn't give too much thought into it.
She would like to show a certain side of her to Robbie and the outside world.
\end{snippetexercise}

\begin{snippetexercise}{atonement-ex-6}%
{Why is the passage below significant? What does the letter represent to Briony?}
\hspace{2cm}\makecell[l]{
    \textit{With the letter, something elemental, brutal, perhaps even criminal had been} \\
    \textit{introduced, some principle of darkness, and even in her excitement over the} \\
    \textit{possibilities, she did not doubt that her sister was in some way} \\
    \textit{threatened and would need her help.}
}

Briony is too young to understand the contents of the letters and
misinterprets it as a threat and wants to protect her sister.
Briony assumes that she has now become the protector of her sister. 
\end{snippetexercise}

\begin{snippetexercise}{atonement-ex-7}%
{Why does she react so strongly to reading \quotes{the word} on the letter.}
Briony never read that word, no one in her presence had ever referred to it.
She cannot even directly refer to the word as it disgusted her profoundly,
she cannot handle what the word means as she is too young.
Lola also adds to the crazy fantasy that Briony has created in her head
as she also infers a threat from the word, since she is also not mature enough.
\end{snippetexercise}

\begin{snippetexercise}{atonement-ex-8}%
{How does chapter 10 end?}
The scenes ends through the eyes of Briony as she watches Robbie and Cecilia
having sex in the library. Briony misinterprets the acts and thinking Robbie is assaulting her.
\end{snippetexercise}

\end{document}
