\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bettelini}
\usepackage{stellar}
\usepackage{makecell}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={English},
    pdfpagemode=FullScreen,
}

\begin{document}

\id{english-atonement-chapter4-7}
\genpage

\section{Summary}

\section{Quotes}

\begin{snippet}{atonement-chapter4-7-quotes}
\begin{minipage}[l]{0.05\textwidth}
    \circled{1}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Quote:} \textit{
        IT WAS not until the late afternoon that
        Cecilia judged the vase repaired. It had
        baked all afternoon on a table by a south-
        facing window in the library, and now three
        fine meandering lines […] were all that
        showed. No one would ever know. (p. 40)
    }
    \\
    \textbf{Who speaks / Who sees?:}
    The narrator is third person but we see through Cecilia's eyes.
    Thus, she is the focalizer.
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{2}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Quote:} \textit{
        [S]he didn't wish to dirty her cashmere by
        lying on the floor, and instead slumped in a
        chair, and the director could hardly object to
        that. The older girl entered so fully into the
        spirit of her own aloof compliance that she
        felt beyond reproach. (p. 52)
    }
    \\
    \textbf{Who speaks / Who sees?:}
    The action is being perceived and judged by Lola (internal focalizer),
    but the narrator is third person.
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{3}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Quote:} \textit{
        NOT LONG after lunch, once she was assured
        that her sister's children and Briony had
        eaten sensibly and would keep their promise
        to stay away from the pool for at least two
        hours, Emily Tallis had withdrawn from the
        white glare of the afternoon's heat to a cool
        and darkened bedroom. (p. 60)
    }
    \\
    \textbf{Who speaks / Who sees?:}
    There is a third person narrator but the internal focalizer is
    internal (Emily).
\end{minipage}
\hr
\begin{minipage}[l]{0.05\textwidth}
    \circled{4}
\end{minipage}
\begin{minipage}[r]{0.95\textwidth}
    \textbf{Quote:} \textit{
        In a spirit of mutinous resistance, she
        climbed the steep grassy slope to the bridge,
        and when she stood on the driveway, she
        decided she would stay there and wait until
        something significant happened to her. (p.
        72)
    }
    \\
    \textbf{Who speaks / Who sees?:}
    There is a third person narrator but the internal focalizer is Briony.
\end{minipage}
\end{snippet}

\plain{All of these are third person narrator, but the internal focalizer
changes constantly between characters.}

\section{Exercises}

\begin{snippetexercise}{atonement-ex-4}%
{For each item taken from this section of the novel, describe what it is and why it's relevant to the
story}
\begin{itemize}
    \item \textbf{\quotes{the look}}
        It's about Ccilia and Leon, that can communicate by looking at each other
        without others understanding them. They share a strong bond and in this case
        Leon wishes Cecilia and Paul to marry, but she's not really interested.
    \item \textbf{cigarettes}
        Cecilia smokes even thought is it not very lady-like.
        Cigarettes are sometimes related to feminism,
        indicating a rebellion against society's standards.
    \item \textbf{divorce}
        The parents of the twins and Lola are divorcing. Divorce was
        very rare and taboo and talking about it is quite forbidden.
        The children are spent to the Tallis family to spend time there
        for a while.
    \item \textbf{Hamlet}
        Hamlet is mentioned by Paul Marshall when talking to Lola.
        Both of them pretends to have seen it even thought they really haven't.
    \item \textbf{Army Amo}
        This is a pun because \textit{amo} is the latin word for love,
        but amo is also similar to ammo - this double meaning between
        ammunition and love is the surname given to Paul's chocolate,
        as a slong for loving you country and fueling the soldier.
        Lola is told about this flirtatiously.
    \item \textbf{curtains}
        Emily goes in her room and closes the curtains - this is a sign
        to tell the world that she doesn't want be disturbed.
    \item \textbf{nettle branch}
        Nettle branch is a simple branch that Briony uses when she sees that the reherasals
        are not doing well, destroying the branch to show her frustration.
\end{itemize}
\end{snippetexercise}

\end{document}
