\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bettelini}
\usepackage{stellar}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={English},
    pdfpagemode=FullScreen,
}

\begin{document}

\title{English}
\id{english-frankenstein-chapter-15-16}
\genpage

\subsection{Exercises}

\begin{snippetexercise}{frankenstein-ex-27}
    {What impact do the three books the creature finds have on him?}
    The books that the creature finds help him understand new things
    about humanity and himself. He is excited and depressed at the same time because
    he started to acknowledge that he's different from everyone else.
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-28}
    {The creature is particularly struck by John Milton's Paradise Lost, whose contents teach him about his
    creation. How, in his opinion, is the monster different from Adam, and yet also different from Satan?}
    Adam, just like the monster, is created unnaturally but the difference
    between them is that Adam was created by God and he was perfect. 
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-29}
    {Why does the creature introduce himself to old De Lacey only when his children and the servants are not
    at home?}
    The creature introduces himself to the old man because he is blind; he cannot see him
    for what he is.
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-30}
    {How does the creature feel as a result the encounter with the cottagers?}
    His angel expanded to the whole human kind.
    If they don't accept him, who could? Revenge feelings
    arise in him. 
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-31}
    {What happens on the boarder to Switzerland? What impact does this incident have on the creature?}
    He helps a little girl but a person sees him and shoots him.
    Hate and revenge reach the peak in the monster's feelings.
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-32}
    {What happens at the end of chapter 16?}
    He kills William and makes sure someone else gets the blame.
\end{snippetexercise}

\begin{snippetexercise}{frankenstein-ex-21}
    {What does the creature want from Frankenstein?}
    He wants a creature/companion just like him.
\end{snippetexercise}

\end{document}
