\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stellar}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Stellar},
    pdfpagemode=FullScreen,
}

\begin{document}

\id{fisica-fluidostatica}
\genpage

\section{Hydrostatic Pressure}

\begin{snippetdefinition}{hydrostatic-pressure-definition}{Hydrostatic pressure}
    The hydrostatic pressure is given by
    
    \begin{align*}
        P &= \frac{F_g}{A} = \frac{A \cdot h \cdot \rho \cdot g}{A} \\
        &= \rho \cdot h \cdot g
    \end{align*}
\end{snippetdefinition}

\plain{This means that the pressure of a fluid doesn't depend
on the area.}

\section{Stevin's law}

\begin{snippetdefinition}{stevin-law-definition}{Stevin's law}
    Stevin's law states that the pressure at any point
    in a resting fluid is only proportional to the depth of that point.

    \[
        P = P_0 + \rho \cdot g \cdot h
    \]

    where \(P_0\) is the external pressure (pressure at the surface).
\end{snippetdefinition}

\section{Pascal's Principle}

\begin{snippetdefinition}{pascal-principle-definition}{Pascal's Principle}
    Pascal's Principle states that a pressure change in one part
is transmitted without loss to every portion of the fluid and walls of the container.
\end{snippetdefinition}

\section{Archimede's Law}

\subsection{Definition}

\begin{snippetdefinition}{archimede-law-definition}{Archimede's law}
    Archimedes' principle states that
    everybody partially or completely immersed in a fluid receives a
    vertical thrust from the bottom upwards,
    equal in intensity to the weight of the displaced fluid.

    The intensity of the force is given by

    \[
        F = \rho_\text{fluid} \cdot g \cdot V_\text{body}
    \]

    The force is the same no matter the shape of the body.
\end{snippetdefinition}

\subsection{Accelleration}

\begin{snippet}{body-in-fluid-acceleration}
    An object dropped in a fluid will sink or float up
    depending on its density and the density of the fluid.
    The body will move with the following acceleration

    \begin{align*}
        F &= F_g - F_\text{Archimede} \\
        &= gm - \rho_\text{fluid}V_\text{body}g \\
        &= gm - \rho_\text{fluid}\frac{m}{\rho_\text{body}}g \\
        ma &= m \left( g- \frac{\rho_\text{fluid}g}{\rho_\text{body}} \right) \\
        a &= \frac{g\rho_\text{body}}{\rho_\text{body}} - \frac{g\rho_\text{fluid}}{\rho_\text{body}} \\
        &= \frac{\rho_\text{body} - \rho_\text{fluid}}{\rho_\text{body}}g
    \end{align*}

    where \(m\) is the mass of the body.

    A submerged body will

    \begin{align*}
        \begin{cases}
            \text{if } \rho_\text{body} < \rho_\text{fluid},\quad \text{float up} \\
            \text{if } \rho_\text{body} = \rho_\text{fluid},\quad \text{float} \\
            \text{if } \rho_\text{body} > \rho_\text{fluid},\quad \text{sink}
        \end{cases}
    \end{align*}
    \phantom{}
\end{snippet}

\end{document}