\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stellar}
\usepackage{enumitem}
\usepackage{definitions}

\begin{document}

\id{linearalgebra-exercises-batch-3}
\genpage

\section{Exercises}

\begin{snippetexercise}{linear-algebra-batch-3-ex-1}{}
    On the \set \(W = \realnumbers^2\) consider the following operations of
    addition and multiplication by a real scalar:
    \((x_1, y_1) \oplus (x_2, y_2) \triangleq (x_1 + x_2 + 1, y_1 + y_2 + 1)\),
    for all \((x_1, y_1), (x_2, y_2) \in W\); \(\lambda \odot (x, y) \triangleq (\lambda x, \lambda y)\),
    for all \(\lambda \in \realnumbers\) and for all \((x, y) \in W\).
    Prove that \((W, \oplus, \odot)\) is not a vector space.
    Which axioms are not verified?
\end{snippetexercise}

\begin{snippetsolution}{linear-algebra-batch-3-ex-1-proof}{}
    Let us check the distributive property of scalar multiplication over vector addition:
    \(\lambda \odot (v \oplus w) = (\lambda \odot v) \oplus (\lambda \odot w)\).
    Let \(v = (x_1, y_1)\) and \(w = (x_2, y_2)\).
    On the left side we have:
    \begin{align*}
        \lambda \odot (v \oplus w) &= \lambda \odot (x_1 + x_2 + 1, y_1 + y_2 + 1) \\
        &= (\lambda(x_1 + x_2 + 1), \lambda(y_1 + y_2 + 1)) \\
        &= (\lambda x_1 + \lambda x_2 + \lambda, \lambda y_1 + \lambda y_2 + \lambda)
    \end{align*}
    On the right side we have:
    \begin{align*}
        (\lambda \odot v) \oplus (\lambda \odot w) &= (\lambda x_1, \lambda y_1) \oplus (\lambda x_2, \lambda y_2) \\
        &= (\lambda x_1 + \lambda x_2 + 1, \lambda y_1 + \lambda y_2 + 1)
    \end{align*}
    These two are equal if and only if \(\lambda = 1\), which must instead hold for all \(\lambda \in \realnumbers\).
    Thus, the distributive property is not verified, and \((W, \oplus, \odot)\) is not a vector space.
\end{snippetsolution}

\begin{snippetexercise}{linear-algebra-batch-3-ex-2}{}
    Let \(A \in \matrices_{m\times n}(\mathbb{R})\) and \(b\in \mathbb{F}^m\),
    study when
    \[
        \text{Sol}(Ax = b) = \{
            x \in \mathbb{F}^n \suchthat Ax = b
        \}
    \]
    is a linear subspace of \(\mathbb{F}^n\).
\end{snippetexercise}

\begin{snippetsolution}{linear-algebra-batch-3-ex-2-proof}{}
    We need
    \[
        0 \in \text{Sol}(Ax = b)
    \]
    but this can only happen if \(b = 0\). So the \set
    is a linear subspace of \(\mathbb{F}^n\) \ifandonlyif
    \(b=0\).
\end{snippetsolution}

\begin{snippetexercise}{linear-algebra-batch-3-ex-3}{}
    Determine which of the following subsets of \(\realnumbers^3\) are vector subspaces of \(\realnumbers^3\):\\
    \begin{enumerate}[label=\Alph*]
        \item \(X_1 = \{(x_1, x_2, x_3) \in \realnumbers^3 \suchthat x_1 + x_2 + x_3 = 0\}\);
        \item \(X_2 = \{(x_1, x_2, x_3) \in \realnumbers^3 \suchthat x_1^2 + x_2^2 + x_3^2 = 0\}\);
        \item \(X_3 = \{(x_1, x_2, x_3) \in \realnumbers^3 \suchthat x_1^2 + x_2^2 + x_3^2 = 1\}\);
        \item \(X_4 = \{(x_1, x_2, x_3) \in \realnumbers^3 \suchthat x_1 = x_2 + x_3\}\);
        \item \(X_5 = \{(x_1, x_2, x_3) \in \realnumbers^3 \suchthat (x_1 - 3x_2 + 4x_3)^2 + (x_1 - x_3)^2 + 1 = 0\}\).
    \end{enumerate}
\end{snippetexercise}

\begin{snippetsolution}{linear-algebra-batch-3-ex-3-proof}{}
    \begin{enumerate}[label=\Alph*]
        \item Yes, \(X_1\) is a vector subspace. It is the set of solutions to a homogeneous linear equation.
        \( (0,0,0) \in X_1 \) and it is closed under addition and scalar multiplication.
        \item Yes, \(X_2\) is a vector subspace. Since \(x_1,x_2,x_3 \in \realnumbers\), \(x_1^2+x_2^2+x_3^2=0 \implies x_1=x_2=x_3=0\). Thus \(X_2 = \{(0,0,0)\}\), which is the trivial vector subspace.
        \item No, \(X_3\) is not a vector subspace since \((0,0,0) \notin X_3\) (\(0^2+0^2+0^2 = 0 \neq 1\)).
        \item Yes, \(X_4\) is a vector subspace. It can be rewritten as \(x_1 - x_2 - x_3 = 0\), which is a homogeneous linear equation like in A.
        \item Yes, \(X_5\) is a vector subspace. Since squares of real numbers are non-negative, the sum \((x_1-3x_2+4x_3)^2 + (x_1-x_3)^2 = -1\) has no real solutions. Thus \(X_5 = \emptyset\). Strictly speaking, the empty set is not considered a vector space because a vector space must contain the zero vector. So \(X_5\) is not a vector subspace.
    \end{enumerate}
\end{snippetsolution}

\begin{snippetexercise}{linear-algebra-batch-3-ex-4}{}
    By convention, the degree of the zero polynomial is \(-\infty\). 
    Given \(m \in \integers_{\geq 0}\),
    let
    \begin{align*}
        V &\triangleq \{p \in \mathcal{P}(\mathbb{F}) \suchthat \polynomialdeg(p) \leq m\} \\
        W &\triangleq \{p \in \mathcal{P}(\mathbb{F}) \suchthat \polynomialdeg(p) = m\} \\
        Z &\triangleq \{p \in \mathcal{P}(\mathbb{F}) \suchthat \polynomialdeg(p) \geq m\}
    \end{align*}
    Determine whether \(V\), \(W\) or \(Z\) are subspaces of \(\mathcal{P}(\mathbb{F})\).
\end{snippetexercise}

\begin{snippetsolution}{linear-algebra-batch-3-ex-4-proof}{}
    \begin{itemize}
        \item \(V\) is a subspace. The zero polynomial has degree \(-\infty \leq m\), so \(0 \in V\).
        If \(p, q \in V\), then \(\polynomialdeg(p+q) \leq \max(\polynomialdeg(p), \polynomialdeg(q)) \leq m\), so \(p+q \in V\).
        If \(p \in V\) and \(\lambda \in \mathbb{F}\), \(\polynomialdeg(\lambda p) \leq \polynomialdeg(p) \leq m\), so \(\lambda p \in V\).
        \item \(W\) is not a subspace. The zero polynomial has degree \(-\infty \neq m\), so \(0 \notin W\).
        Furthermore, it's not closed under addition (e.g. \((x^m) + (-x^m) = 0 \notin W\)).
        \item \(Z\) is not a subspace. The zero polynomial has degree \(-\infty \ngeq m\), so \(0 \notin Z\). 
        Like \(W\), it is not closed under addition.
    \end{itemize}
\end{snippetsolution}

\end{document}