\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stellar}
\usepackage{definitions}

\newcommand\ts{{(X, \mathcal{T})}}

\begin{document}

\id{topology-connectedness}
\genpage

\section{Connectedness}

\begin{snippetdefinition}{connected-topological-space-definition}{Connected topological space}
    Let \(X\) be a \topologicalspace.
    Then, \(X\) is said to be \emph{connected} if the only subsets that are
    both \openset[open] and \closedset[closed]
    are \(\emptyset\) and \(X\).
\end{snippetdefinition}

\plain{If there were other such sets, then the space would be homeomorphic to the disjoint union of the set and its complement.}

\begin{snippetlemma}{connected-topological-space-equivalence}{}
    Let \(X\) be a \topologicalspace.
    Then, \(X\) is not connected \ifandonlyif it is disjoint union
    of two non-empty \openset[open sets] (or \closedset[closed sets]).
\end{snippetlemma}

\begin{snippetproposition}{connected-topological-space-subspace}{}
    Let \(X\) be a \topologicalspace and \(Y \subseteq X\) and \(Y \neq \emptyset\) be a connected subspace of \(X\).
    Then, for every \(A \subseteq X\) that is both \openset[open] and \closedset[closed],
    we have
    \[
        Y \subseteq A \lxor Y \intersection Y = \emptyset
    \]
    If \(Y = \emptyset\) they are both true.
\end{snippetproposition}

\begin{snippetproof}{connected-topological-space-subspace-proof}{connected-topological-space-subspace}{}
    Since \(A\) is both \openset[open] and \closedset[closed],
    \(A \intersection Y\) is \openset[open] and \closedset[closed] in \(Y\).
    Since \(Y\) is connected, \(A \intersection Y = \emptyset\) or \(A \intersection Y = Y\)
\end{snippetproof}

\begin{snippetproposition}{connectedness-of-real-unital-interval}{Connectedness of real unital interval}
    Consider \([0,1] \subseteq \realnumbers\) with the subspace topology
    induced by the standard real topology.
    Then, \([0,1]\) is connected.
\end{snippetproposition}

\begin{snippetproof}{connectedness-of-real-unital-interval-proof}{connectedness-of-real-unital-interval}{Connectedness of real unital interval}
    Let \(C, D \subseteq [0,1]\)
    be non-empty \closedset[closed] subspaces of \([0,1]\)
    such that \(C \union D = [0,1]\).
    They are also \closedset[closed] in \(\realnumbers\).
    We want to show that \(C \intersection D \neq \emptyset\).
    Let \(d = \inf D\), which exists since \(D \subseteq [0,1]\).
    If \(d=0\), \(C \intersection D \neq \emptyset\).
    Otherwise, let \(d \neq 0\). Then,
    \([0, d) \subseteq C\). This is given by the fact that \(C \union D = [0,1]\)
    and \(d = \inf D\).
    Since \(C\) is \closedset[closed], \(d \in \overline{[0,d)} \subseteq \overline{C} = C\).
    On the other hand, \(d \in D\) by definition of \(\inf D\).
    Thus \(d \in \overline{D} = D\).
    Finally, \(c\in C \intersection D\) and thus \(C \intersection D \neq \emptyset\).
\end{snippetproof}

\begin{snippetproposition}{continuous-map-preserves-connectedness}{}
    Let \(X,Y\) be \topologicalspace[topological spaces]
    such that \(X\) is connected
    and \(f \colon X \fromto Y\) be a map.
    Then, \(f(X)\) is connected.
\end{snippetproposition}

\begin{snippetproof}{continuous-map-preserves-connectedness-proof}{continuous-map-preserves-connectedness}{}
    Let \(Z \subseteq f(X)\)
    be a non-empty subset of \(f(X)\), \openset[open]
    and \closedset[closed] in \(f(X)\). We want to show that \(Z = f(X)\).
    We have \(Z = A \intersection f(X) = C \intersection f(X)\)
    for some \openset \(A\) of \(Y\)
    and some \openset \(C\) of \(Y\).
    Then, \(f^{-1}(Z) = f^{-1}(A) = f^{-1}(C)\).
    Note that \(f^{-1}(Z) \neq \emptyset\).
    Indeed, \(Z = f(f^{-1(Z)})\)
    and \(Z \subseteq f(X)\).
    This is true because we restricted the codomain of \(f\) and it is thus \surjective.
    If it were \(f^{-1}(Z) = \emptyset\) we would have \(Z = f(\emptyset) = \emptyset\),
    which would be absurd \lightning.
    Now, since \(X\) is connected, it follows that
    \[
        f^{-1}(Z) = X
    \]
    But then \(f(X) = f(f^{-1}(Z)) = Z\).
\end{snippetproof}

\subsection{Arc connectedness}

\begin{snippetdefinition}{topological-path-definition}{Topological path}
    Let \(X\) be a \topologicalspace.
    A continuous map \(\varphi \colon [0,1] \fromto X\)
    is said to be a \emph{path}.
\end{snippetdefinition}

\begin{snippetdefinition}{arc-connected-topological-space-definition}{Arc connected topological space}
    Let \(X\) be a \topologicalspace.
    Then, \(X\) is said to be \emph{arc connected} if \(\forall x,y \in X\)
    there exists a path \(\varphi\colon [0,1] \to X\)
    such that \(\varphi(0)=x\) and \(\varphi(1)=y\).
\end{snippetdefinition}

\begin{snippetproposition}{arc-connectedness-implies-connectedness}{}
    Let \(X\) be a \topologicalspace 
    that is arc connected. Then, \(X\) is connected.
\end{snippetproposition}

\begin{snippetproof}{arc-connectedness-implies-connectedness-proof}{arc-connectedness-implies-connectedness}{}
    Let \(A,B\) be non-empty \openset[open sets] of \(X\)
    such that \(A \union B = X\).
    We want to show that \(A \intersection B \neq \emptyset\).
    We know there \(\exists x \in A\) and \(\exists y \in B\).
    By hypothesis, there exists \(\alpha \colon [0,1] \fromto \realnumbers\)
    such that \(\alpha(0) = x\) and \(\alpha(1) = y\).
    Since \(\alpha\) is continuous by definition,
    \(\alpha^{-1}(A)\) and \(\alpha^{-1}(B)\) are \openset[open]
    in \([0,1]\). We know that \(0 \in \alpha^{-1}(A)\)
    and \(1 \in \alpha^{-1}(B)\).
    These are non-empty \openset[open sets]
    \[
        [0,1] = \alpha^{-1}(X)  = \alpha^{-1}(A \union B)
        = \alpha^{-1}(A) \union \alpha^{-1}(B)
    \]
    Since \([0,1]\) is connected, \(\alpha^{-1}(A) \intersection \alpha^{-1}(B) \neq \emptyset\).
    Thus, \(\exists t \in \alpha^{-1}(A) \intersection \alpha^{-1}(B)\)
    which means \(\alpha(t) \in A \intersection B \neq \emptyset\).
\end{snippetproof}

\begin{snippetlemma}{topological-subspace-disjoint-sets-arc-connectedness}{}
    Let \(A, B\) topological subspaces of a \topologicalspace \(X\)
    such that \(A\) and \(B\) are arc connected.
    Suppose \(A \intersection B \neq \emptyset\).
    Then, \(A \union B\) is arc connected.
\end{snippetlemma}

\begin{snippetproof}{topological-subspace-disjoint-sets-arc-connectedness-proof}{topological-subspace-disjoint-sets-arc-connectedness}{}
    Given \(x,y \in A \union B\) we need to show theat there exists a path between them.
    If \(x,y \in A\) or \(x,y \in B\) the thesis follows by the hypothesis.
    Suppose without loss of generality that \(x\in A \land y \in B\),
    consider a point \(z \in A \intersection B\).
    Since \(A\) is arc connected there exists a path \(\alpha \colon [0,1] \fromto X\)
    such that \(\alpha(0) = x\) and \(\alpha(1) = z\).
    Since \(B\) is arc connected there exists a path \(\beta \colon [0,1] \fromto X\)
    such that \(\beta(0) = z\) and \(\beta(1) = y\).
    Then, we create the path \(\gamma \colon [0,1] \fromto X\)
    defined by \[
        \gamma = \alpha \ast \beta = \begin{cases}
            \alpha(2t) & t \leq \frac{1}{2} \\
            \beta(2t-1) & t > \frac{1}{2}
        \end{cases}
    \]
    Then,
    \(\gamma(0) = x\) and \(\gamma(1) = y\).
\end{snippetproof}

\begin{snippetproposition}{sn-is-arc-connected}{}
    \(S^n\) is arc connected.
\end{snippetproposition}

\begin{snippetproof}{sn-is-arc-connected-proof}{sn-is-arc-connected}{}
    Consider \(S^n \difference \{p\}\) and consider the stereographical projection.
    This is a homeomorphism with \(\realnumbers^n\) since \(S^n \subset \realnumbers^{n+1}\).
    Since we can choose two distinct points \(a, b\) we have
    \[
        S^n = (S^n - \{a\}) \union (S^n - \{b\})
    \]
    Both these terms are arc connected and by \snippetref[topological-subspace-disjoint-sets-arc-connectedness][this lemma],
    \(S^n\) is connected.
\end{snippetproof}

\plain{Every convex real hyperdimensional subset is arc connected.}

\subsection{Connected components}

\begin{snippetdefinition}{connected-component-definition}{Connected component}
    Let \(X\) be a \topologicalspace.
    A \emph{connected component} of \(X\) is a maximal connected subspace of \(X\).
    That is, a non-empty connected subspace \(C\) such that
    for any connected subspace \(A \supseteq C\), we have \(A = C\).
\end{snippetdefinition}

\begin{snippetlemma}{closure-of-connected-is-connected}{Closure of connected}
    Let \(Y\) be a connected subspace of a \topologicalspace \(X\).
    Let \(W\) be a subspace such that \(Y \subseteq W \subseteq \closure{Y}\).
    Then \(W\) is connected.
\end{snippetlemma}

\begin{snippetproof}{closure-of-connected-is-connected-proof}{closure-of-connected-is-connected}{Closure of connected}
    Let \(Z \subseteq W\) be \openset[open], \closedset[closed], and non-empty.
    Consider \(Z \intersection Y\), which is open and closed in \(Y\)
    by transitivity of the subspace topology.
    
    Since \(Z \neq \emptyset\) is open in \(W\),
    we have \(Z = A \intersection W\) for some open \(A\) in \(X\).
    For any \(x \in Z \subseteq A\), by definition of closure we have
    \(A \intersection Y \neq \emptyset\).
    
    Since \(Y\) is connected and \(Z \intersection Y\) is open, closed, and non-empty in \(Y\),
    we deduce \(Z \intersection Y = Y\), i.e., \(Y \subseteq Z\).
    
    Taking closures in \(W\): \(\overline{Y} \subseteq \overline{Z} = Z\)
    (since \(Z\) is closed in \(W\)).
    Since \(W \subseteq \overline{Y}\), we get \(W \subseteq Z\), hence \(Z = W\).
\end{snippetproof}

\begin{snippetlemma}{union-of-connected-with-common-point}{Union of connected sets}
    Let \(X\) be a \topologicalspace, \(x \in X\), and
    \(\{Z_i\}_{i \in I}\) a family of connected subspaces such that
    \(x \in Z_i\) for all \(i\).
    Then \(\bigcup_{i \in I} Z_i\) is connected.
\end{snippetlemma}

\begin{snippetproof}{union-of-connected-with-common-point-proof}{union-of-connected-with-common-point}{Union of connected sets}
    Let \(W = \bigcup_{i \in I} Z_i\) and let \(A \subseteq W\)
    be open, closed, and non-empty.
    
    For each \(i\), consider \(A \intersection Z_i\) which is open and closed in \(Z_i\).
    Since \(Z_i\) is connected, either \(A \intersection Z_i = \emptyset\)
    or \(A \intersection Z_i = Z_i\) (i.e., \(Z_i \subseteq A\)).
    
    Since \(A \neq \emptyset\), there exists some \(i_0\) with \(A \intersection Z_{i_0} \neq \emptyset\),
    hence \(Z_{i_0} \subseteq A\), so \(x \in A\).
    
    But then for every \(i\), we have \(x \in Z_i \intersection A \neq \emptyset\),
    hence \(Z_i \subseteq A\) for all \(i\).
    Therefore \(W \subseteq A\), so \(A = W\).
\end{snippetproof}

\begin{snippetcorollary}{union-two-connected-with-intersection}{Union of two connected sets}
    Let \(A, B\) be connected subspaces of a \topologicalspace.
    If \(A \intersection B \neq \emptyset\), then \(A \union B\) is connected.
\end{snippetcorollary}

\begin{snippettheorem}{space-is-union-of-components}{Decomposition into components}
    Every \topologicalspace \(X\) is the union of its connected components.
    Each connected component is \closedset[closed],
    and every point belongs to exactly one connected component.
\end{snippettheorem}

\begin{snippetproof}{space-is-union-of-components-proof}{space-is-union-of-components}{Decomposition into components}
    For each \(x \in X\), define \(C(x)\) as the union of all connected subspaces
    containing \(x\). By the previous lemma, \(C(x)\) is connected.
    
    Since \(x \in C(x)\), we have \(X = \bigcup_{x \in X} C(x)\).
    
    By the closure lemma, \(\overline{C(x)}\) is connected.
    Since \(C(x)\) is maximal connected, \(C(x) = \overline{C(x)}\), so \(C(x)\) is closed.
    
    If \(C, D\) are components with \(C \intersection D \neq \emptyset\),
    then \(C \union D\) is connected.
    By maximality, \(C = C \union D = D\).
\end{snippetproof}

\begin{snippetdefinition}{pi-zero-definition}{\(\pi_0\)}
    Let \(X\) be a \topologicalspace.
    We define \(\pi_0(X)\) as the set of connected components of \(X\).
\end{snippetdefinition}

\end{document}