\documentclass[preview]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{stellar}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{blkarray}
\usepackage{tikz}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Graphs},
    pdfpagemode=FullScreen,
}

\begin{document}

\title{Graphs}
\id{graphs}
\genpage

\section{Definition}

\begin{snippetdefinition}{graph-definition}{Graph}{
    A \textit{graph} is a tuple \((V,E)\) consisting of a set of \textit{vertices} \(V\)
    and a set of \textit{edges} \(E\) where every element in \(E\) is a distinct pair of vertices
    in \(V\).
}
\end{snippetdefinition}

\section{Degree}

\begin{snippetdefinition}{graph-vertex-degree}{Graph}{
    Let \(G=(V,E)\) be a graph. The \textit{degree} of a vertex \(v \in V\), denoted \(deg(v)\) is defined
    as the numbers of edges in \(E\) that are incident on \(v\).
}
\end{snippetdefinition}

\begin{snippetproposition}{graph-sum-of-degrees}{Sum of degrees of graph}{
    Let \(G=(V,E)\) be a graph.
    The sum of the degrees of all vertices is equals to twice the number of edges.
    \[
        \sum_{v\in V}deg(v)=2|E|
    \]
}
\end{snippetproposition}

\section{Paths}

\begin{snippetdefinition}{path-definition}{Path}{
    Let \(G=(V, E)\) be a graph.
    A \textit{path} on \(G\) is a sequence of vertices in \(V\) connected by edges in \(E\).
}
\end{snippetdefinition}

\begin{snippetdefinition}{ciclic-path-definition}{Cycle}{
    Let \(P\) be a path on a graph \(G=(V, E)\).
    The path is said to be \textit{cycle} if it starts and end at the same vertex.
}
\end{snippetdefinition}

\begin{snippetdefinition}{simple-path-definition}{Simple path}{
    Let \(P\) be a path on a graph \(G=(V, E)\).
    The path is said to be \textit{simple} if every vertex in it is distinct.
}
\end{snippetdefinition}

\begin{snippetdefinition}{connected-path-definition}{Connected path}{
    Let \(G=(V, E)\) be a graph.
    The graph is said to be \textit{connected} if there exist
    at least a path between any pair of vertices.
}
\end{snippetdefinition}

\section{Adjacency Matrices}

\begin{snippetdefinition}{adjacency-matrix-definition}{Adjacency Matrix}
    A finite graph can be represented by a square matrix
    \(n \times n\) where \(n\) is the number of vertices.

    \begin{center}
        \begin{minipage}[r]{6cm}
            \begin{tikzpicture}[main/.style = {draw, circle}]
                \foreach \x [count = \xi] in {a,b,c,d,e} {
                    \node[main] (\xi) at
                        ({1 * cos(360 / 5 * \xi)}, {1 * sin(360 / 5 * \xi)})
                        {\(\x\)};
                }
            
                \draw (1) -- (2);
                \draw (1) -- (3);
                \draw (1) -- (5);
                \draw (2) -- (4);
                \draw (5) -- (3);
            \end{tikzpicture} 
        \end{minipage}
        \begin{minipage}[l]{6cm}
            A=
            \begin{blockarray}{cccccc}
                & a & b & c & d & e \\
                \begin{block}{c(ccccc)}
                a & 0 & 1 & 1 & 0 & 1 \\
                b & 1 & 0 & 0 & 1 & 0 \\
                c & 1 & 0 & 0 & 0 & 1 \\
                d & 0 & 1 & 0 & 0 & 0 \\
                e & 1 & 0 & 1 & 0 & 0 \\
                \end{block}
            \end{blockarray}
        \end{minipage}
    \end{center}

    Every row and column represents a vertex. \(1\) means
    that the two vertices are adjacent, \(0\) otherwise. The diagonal of this matrix
    will always e \(0s\) since no vertice is adjacent to itself
    and \(A=A^t\)
\end{snippetdefinition}

\end{document}
