\documentclass[a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{bussproofs}

\usepackage{stackengine}

\newcommand\vdashsub[1]{\mathrel{\stackengine{.4ex}{\vdash}{\;\;\scriptscriptstyle#1}{U}{c}{F}{T}{L}}}

\author{Paolo Bettelini}
\date{}

\begin{document}

\maketitle
\tableofcontents
\pagebreak

\section{Boolean Algebra}

% Associativity
% Commutativity
% Distribuibity

\begin{align*}
    x \lor 0 = x \\
    x \land 0 = 0 \\
    x \lor 1 = 1 \\
    x \land 1 = x \\
    x \lor x = x \\
    x \land x = x \\
    x \land (x \lor y) = x \\
    x \lor (x \land y) = x \\
    x \land \lnot x = 0 \\
    x \lor \lnot x = 1 \\
    \lnot x \land \lnot y = \lnot (x \lor y) \\
    \lnot x \lor \lnot y = \lnot (x \land y)
\end{align*}

\pagebreak

\section{Induction}

Induction can be used to prove a statement in the form \(P(n)\)
for all \(n \in \naturalnumbers\).
For induction to work one must need to prove the base base, which is usually \(P(0)\),
but other starting points can be used to prove the statement from that point on.

Weak and strong induction are equivalent.

\subsection{Weak induction}

Proving that the base case \(P(0)\) is true,
along with proving the induction step \(P(n) \implies P(n+1)\), implies \(P(n)\)
for all \(n \geq 0\). In second-order logic
\[
    \forall P \left(
        P(0) \land \forall n \left( P(n) \implies P(n+1) \right)
        \implies \forall n \left( P(n) \right)
    \right)
\]

\subsection{Strong induction}

Proving that the base case \(P(k)\) for \(k < m\) is true, along with
proving the induction step \(P(m)\), implies \(P(n)\)
for all \(n\).

The base case here is when we need to prove \(P(m)\) for \(m=0\).

\pagebreak

\section{Propositional Logic}

\subsection{\(k\)-ary Boolean function}

A \(k\)-ary Boolean function is a mapping from \({\{T, F\}}^k \to \{T,F\}\)

\subsection{\(0\)-ary Boolean function}

The \(0\)-ary Boolean function are the \textit{verum} (\(\top\)) and \textit{falsum} (\(\bot\)) connectives.
The represent respectively the True value and the False value.

\subsection{Propositional formula}

A \textit{propositional formula} is a formula which has a unique truth value given all variables.

The set of all formulas is countable.

\subsection{Truth assignment}

A \textit{truth assignment} is a function which maps a set of propositional
variables \(V=\{p_1, p_2, \ldots, p_n\}\) to a boolean value
\[
    \tau: V \to \{T,F\}
\]

A formula \(A\) involving the variables \(V=\{p_1, p_2, \ldots, p_n\}\)
defines a \(k\)-ary boolean function \(f_A(x_1, x_2, \ldots, x_n)\) where \(x_n = \tau(p_n)\).


%%%%%%%%%%% https://www.youtube.com/watch?v=b_FLLGTjreM&list=WL&index=1&t=1052s

%\subsection{Language}
%
%A \textit{language} \(L\) is a set of connectives which may be used to describe
%an \(L\)-formula.
%
%A language \(L\) is \textit{complete} iff every \(k\)-ary boolean functions can be
%defined by an \(L\)-formula.
%
\subsection{Tautology}

A propositional formula \(A\) is a \textit{tautology} \(\vDash A\) if its \(k\)-ary boolean
function \(f_A\) is always \(T\). \\
Otherwise, we say \(\nvDash A\).

\subsection{Satisfiability}

A propositional formula \(A\) is \textit{satisfiable}
if \(f_A\) is \(T\) for some input.

If \(\Gamma\) is a \underline{set} of propositional formulas, \(\Gamma\)
is satisfiable if there are some assignments to satisfy all its members.

\(\Gamma \vDash A\) (tautologically implies \(A\))
if every truth assignment satisfying \(\Gamma\)
also satisfies \(A\).

\subsection{Substitution}

A \textit{substitution} \(\sigma\)
is a mapping from a set of propositional variables
to the set of propositional formulas.
If \(A\) is a propositional formula, \(A\sigma\)
is equal to the formula obtained by simultaneously
replacing each variable appearing in \(A\)
by its image under \(\sigma\).

\subsection{Propositional Proof System}

A \textit{Propositional Proof System}
\(\mathcal{F}\) has every substitution into the
axioms scheme as his axioms and a set of inference rules.
\\
If \(A\) has an \(\mathcal{F}\)-proof,
then \(\vdash A\). %or \(\vdashsub{\mathcal{F}}A\).
If the proof needs extra hypothesis \(\Gamma\)
(which may not be tautologies), then \(\Gamma\vdash A\)
.%or \(\Gamma \vdashsub{\mathcal{F}} A\)

\subsection{Soundness \(\mathcal{F}\)}

\(\mathcal{F}\) is sound iff every \(\mathcal{F}\)-formula is logically valid with respect to the semantics of the system. 

\subsection{Completness \(\mathcal{F}\)}

\(\mathcal{F}\) is complete iff
it can prove any valid formula, meaning that
the semantic noion of validity
and the syntactic notion of provability coincide,
and a formula is valid iff it is has an
\(\mathcal{F}\)-proof.

\end{document}