\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{fullpage}
\usepackage{tikz}
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Euler's Formula},
    pdfpagemode=FullScreen,
}

\title{Euler's Formula}
\author{Paolo Bettelini}
\date{}

\begin{document}

\maketitle
\tableofcontents
\pagebreak

\section{Definition}

Euler's formula states that for every \(x\in\mathbb{R}\)
\begin{align*}
	e^{ix}=\cos\,x+i\,\sin\,x
\end{align*}

We can represent the formula on the complex plane

\begin{center}
	\begin{tikzpicture}
		\begin{scope}[thick, font=\scriptsize, scale=4]
			\draw [->] (-1.25,0) -- (1.25,0) node [above left]  {Re(\(s\))};
			\draw [->] (0,-1.25) -- (0,1.25) node [below right] {Im(\(s\))};
		
			\draw (1,-0.75pt) -- (1,0.75pt)   node [above left] {$1$};
			\draw (-1,-0.75pt) -- (-1,0.75pt) node [above left] {$-1$};
			\draw (-0.75pt,1) -- (0.75pt,1)   node [above right] {$i$};
			\draw (-0.75pt,-1) -- (0.75pt,-1) node [below right] {$-i$};

			\path [draw=black,fill=none,semitransparent] (0, 0) circle (1);
			\draw [thick, color=red] (0,0) -- (0.5,0.866);
			\draw [thick, color=blue] (0.5,0) -- (0.5,0.866);
			\draw [thick, color=blue] (0,0) -- (0.5,0);
			
			\draw [color=black, fill=black] (0.5,0.866) circle(0.02) node [above right] {\(e^{i\theta}\)};
			
			\node [below right,gray] at (0.175, -0.05) {\(\cos\,\theta\)};
			\node [below right,gray] at (0.505, 0.43) {\(\sin\,\theta\)};
		\end{scope}
	\end{tikzpicture}
\end{center}

We can notice that \(\left|e^{ix}\right|=1\) since \(|e^{i\theta}|=\cos^2\theta +\sin^2\theta =1\)

\pagebreak

\section{Proof}

To understand this identity we must first look at the Taylor series of some functions.
\\
These functions are all equal to their Taylor series, so we can easily compute their Maclaurin series.

\subsection{Exponential function}

The first thing we can notice is that the derivative of \(f(x)=e^x\) is \(e^x\) itself, so
\begin{align*}
	f^{(n)}(x)=e^x,
	\quad n\in\mathbb{Z}
\end{align*}
which means that for every term of the Maclaurin series
\begin{align*}
	f^{(n)}(a)=f^{(n)}(0)=e^0=1
\end{align*}
The exponential function can then be expressed as the series
\begin{align*}
	e^x=\sum_{n=0}^{\infty}\frac{x^n}{n!}
\end{align*}

\pagebreak

\subsection{Sine function}

Here's a table of the derivatives of \(f(x)=\sin(x)\) and their value at \(a=0\)
\begin{align*}
	\begin{cases}
		f^{(0)}(x)=\sin(x), \quad &f^{(0)}(0)=0 \\
		f^{(1)}(x)=\cos(x), \quad &f^{(1)}(0)=1 \\
		f^{(2)}(x)=-\sin(x),\quad &f^{(2)}(0)=0 \\
		f^{(3)}(x)=-\cos(x),\quad &f^{(3)}(0)=-1\\
	\end{cases}
\end{align*}
The next derivative is \(f^{(4)}(x)=\sin(x)\) making the sequence start over.
\\
We can notice that every even n-th derivative will multiply the term by \(0\).
\\
We will use \(2n+1\) to skip all the even integers.
\\
We can also notice that the sign of the term is given by \({(-1)}^n\) since we only consider odd indexes.
\begin{align*}
	\sin(x)&=x-\frac{x^3}{3!}+\frac{x^5}{5!}-\frac{x^7}{7!}+\cdots
	\\
	&=\sum_{n=0}^{\infty}\frac{x^{2n+1}{(-1)}^n}{(2n+1)!}
\end{align*}

\subsection{Cosine function}

Here's a table of the derivatives of \(f(x)=\cos(x)\) and their value at \(a=0\)
\begin{align*}
	\begin{cases}
		f^{(0)}(x)=\cos(x), \quad &f^{(0)}(0)=1 \\
		f^{(1)}(x)=-\sin(x),\quad &f^{(1)}(0)=0 \\
		f^{(2)}(x)=-\cos(x),\quad &f^{(2)}(0)=-1\\
		f^{(3)}(x)=\sin(x), \quad &f^{(3)}(0)=0 \\
	\end{cases}
\end{align*}
The next derivative is \(f^{(4)}(x)=\cos(x)\) making the sequence start over.
\\
We can notice that every odd n-th derivative will multiply the term by \(0\).
\\
We will use \(2n\) to skip all the odd integers.
\\
We can also notice that the sign of the term is given by \({(-1)}^n\) since we only consider even indexes.
\begin{align*}
	\cos(x)&=1-\frac{x^2}{2!}+\frac{x^4}{4!}-\frac{x^6}{6!}+\cdots
	\\
	&=\sum_{n=0}^{\infty}\frac{x^{2n}{(-1)}^n}{(2n)!}
\end{align*}

\pagebreak

\subsection{Conclusion}

Given the Taylor series for the exponential function
\begin{align*}
	e^x=\sum_{n=0}^{\infty}\frac{x^n}{n!}
\end{align*}
we use \(ix\) instead of \(x\)
\begin{align*}
	e^{ix}=\sum_{n=0}^{\infty}\frac{{(ix)}^n}{n!}
\end{align*}
The imaginary number \(i\) has some amazing properties when it comes to exponentiation.
\begin{align*}
	\begin{cases}
		i^0=+1\\
		i^1=+i\\
		i^2=-1\\
		i^3=-i\\
	\end{cases}
	\quad
	\begin{cases}
		i^4=+1\\
		i^5=+i\\
		i^6=-1\\
		i^7=-i\\
	\end{cases}
	\quad
	\cdots
\end{align*}
We can use these properties to simplify the \(e^{ix}\) Taylor series
\begin{align*}
	e^{ix}
	&   =1
		+ix
		+\frac{{(ix)}^2}{2!}
		+\frac{{(ix)}^3}{3!}
		+\frac{{(ix)}^4}{4!}
		+\frac{{(ix)}^5}{5!}
		+\frac{{(ix)}^6}{6!}
		+\frac{{(ix)}^7}{7!}
		+\cdots
	\\
	\\
	&   =1
		+ix
		-\frac{x^2}{2!}
		-\frac{ix^3}{3!}
		+\frac{x^4}{4!}
		+\frac{ix^5}{5!}
		-\frac{x^6}{6!}
		-\frac{ix^7}{7!}
		+\cdots
	\\
	\\
	&=
	\left(
		1
		-\frac{x^2}{2!}
		+\frac{x^4}{4!}
		-\frac{x^6}{6!}
		+\cdots
	\right)
	+i
	\left(
		x
		-\frac{x^3}{3!}
		+\frac{x^5}{5!}
		-\frac{x^7}{7!}
		+\cdots
	\right)
\end{align*}
We notice that the two terms correspond to the sine and cosine Taylor series
\begin{align*}
	e^{ix}=\cos\,x+i\,\sin\,x
\end{align*}

\pagebreak

\section{Complex trigonometric functions}

Using Euler identify we can define \(\sin(x)\) and \(\cos(x)\) for \(x \in \mathbb{C}\)

\begin{align*}
	\sin(x)&=\frac{e^{ix}+e^{-ix}}{2i} \\
	\cos(x)&=\frac{e^{ix}-e^{-ix}}{2}
\end{align*}

\section{De Moivre's Theorem}

Using the property of exponentiation \(\left(a^b\right)^c = a^{bc}\),
we can see that \(\left(e^{i\theta}\right)^n = e^{in\theta}\).
\\
Using Euler's formula we can deduce that

\[
    \left(\cos(\theta) + i\sin(\theta)\right)^n = \cos(n\theta) + i\sin(n\theta),
	\quad n \in \mathbb{Z}
\]

\section{Nth Roots of Units}

We can extend De Moivre's Theorem for the integers powers or any complex number,
rather than the ones on the unit circle \((r=1)\).

\[
    \left(r\left(\cos(\theta) + i\sin(\theta)\right)\right)^n = 
	r^n\left(\cos(n\theta) + i\sin(n\theta)\right), \quad n \in \mathbb{Z}
\]

The nth roots of \(1\) are the solutions to

\[
	x^n=1
\]

We might write \(1\) as a complex number

\[
	x^n = \cos(0) + i\sin(0)
\]

Comparing this to our extended De Moivre's theorem

\[
	\cos(0) + i\sin(0) = r^n\left(\cos(n\theta) + i\sin(n\theta)\right)
\]

We can see that

\begin{align*}
	r^n&=1 \\
	n\theta&=0
\end{align*}

As long as \(n \neq 0\)

\begin{align*}
	r&=1 \\
	\theta&=0
\end{align*}

\pagebreak

By plugging these values into

\[
	x^n = \left(r\left(\cos(\theta) + i\sin(\theta)\right)\right)^n
\]

we get that \(x=1\).

TODO

% https://youtu.be/Rju2g-sExng?t=1571

\end{document}