\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{fullpage}
\usepackage{hyperref}

\usepackage[version=4]{mhchem}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Radiation},
    pdfpagemode=FullScreen,
}

\title{Radiation}
\author{Paolo Bettelini}
\date{}

\begin{document}

\maketitle
\tableofcontents
\pagebreak

\section{Radioactivity}

\subsection{Definition}

Radioactivity is a set of physical-nuclear processes
through which some unstable or radioactive atomic nuclei decay,
in a certain period of time called decay time.

An unstable nuclei will keep emitting radiations
and transmuting to other nuclei until the atom is stable.

\subsection{Decay}

The mass of a radioactive material will decrease exponentially.

\[
    M(t) = M_0 \cdot e^{-kt}
\]

\(M(t)\) is the mass (or number or particles)
after a certain time \(t\). \(M_0\) is the initial mass
and \(k\) is the rate of decay.

\subsection{Half-life}

The time of half-life is given by \(t_\frac{1}{2} = \frac{\ln 2}{k}\).

\begin{align*}
    \frac{1}{2}M_0 &= M_0 e^{-kt} \\
    \frac{1}{2} &= e^{-kt} \\
    \ln\left(\frac{1}{2}\right) &= -kt \\
    t &= \frac{\ln 2}{k}
\end{align*}

\subsection{Types of radiations}

There are three types of radiations that canbe emitted by an unstable nucleai.

\subsubsection{\alpha{} particles}

An \(\alpha\) particle is a helium nuclei. For example

\[
    \ce{^238_92U -> ^4_2\alpha{} + ^234_90Th}
\]

\subsubsection{\beta{} particles}

There are two types of \(\beta{}\) particles. \(\beta{}^+\) and \(\beta{}^-\).
A \(\beta{}^+\) particle is emitted when the nuclei is unstable due to
having too many protons, whist the \(\beta{}^-\) one is emitted when it has
too many neutrons.

\[
    \begin{cases}
        \beta{}^+,\quad \ce{^0_{+1}e} \text{ (positron)} \\
        \beta{}^-,\quad \ce{^0_{-1}e} \text{ (electron)}
    \end{cases}
\]

\subsubsection{\gamma{} particles}

\(\gamma\) rays are photons of electromagnetic energy. They have \(0\) mass and \(0\) charge.

\end{document}
