\documentclass[a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{fullpage}
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Generic},
    pdfpagemode=FullScreen,
}

\title{Generic}
\author{Paolo Bettelini}
\date{}

\begin{document}


\section{Vector-Valued Function}

A vector-valued function is a function of a real parameter
which returns a vector
\[
    r(t) = \begin{pmatrix}
        f(t) \\
        g(t) \\
        h(t)
    \end{pmatrix}
\]

\section{Tangent Vector Vector-Valued Function}

Given a vector-valued function
\[
    r(t) = \begin{pmatrix}
        f(t) \\
        g(t) \\
        h(t)
    \end{pmatrix}
\]

where \(f\), \(g\) and \(h\) are differentiable,
then the Tangent vector to the curve is given b
\[
    r'(t) = \begin{pmatrix}
        f'(t) \\
        g'(t) \\
        h'(t)
    \end{pmatrix}
\]

\section{Curve length}

The length of the curve beteen a and b is the integral from a to b of \(\sqrt{f'(t)^2+g'(t)^2+h'(t)^2}\).

\section{Curvature}

The tangent vector is given by
\[
    \vec{T} = \frac{d\vec{r}}{ds}
\]

TODO: a smooth curve has \(r'(t)\) continuous and \(r'(t) \neq 0\)

\newcommand\norm[1]{\left\lVert#1\right\rVert}
The curvature of a curve is given by
\[
    \kappa = \norm{\frac{d\vec{T}}{ds}}
\]
where \(\vec{T}\) is the unit tangent and s is the arc length.
By the chain rule this can also be written as
\[
    \norm{\frac{d\vec{T}}{dt}\frac{dt}{ds}} = \frac{1}{\norm{\vec{v}}}\norm{\frac{d\vec{T}}{dt}}
\]
where \(\vec{v}\) is the velocity vector.

% TODO: example, a circle

\section{Principle unit normal}

When \(\kappa \neq 0\)
\[
    \vec{N} = \frac{1}{k}\frac{d\vec{r}}{ds}
\]

The curvature tells how much the curve is curving while the principle
unit normal tells in which direction it is curving (normal to the curvature).

\section{Vectors describing motion}

Given a vector-valued curve \(r(t)\), the motion can be described by
\begin{itemize}
    \item the tangent vector (derivative of the position
        with respect to the arclength parameter), which the direction in which the curve is going.
    \item the normal vector captures the way in which the tangent vector is itself changing.
    \item Binormal vector is the cross product, which represents the vector normal to the plane
        created by the tangent vector and the normal vector. That is, it captures the torsion of said plane.
\end{itemize}

\begin{align*}
    \vec{T} &= \frac{d\vec{r}}{ds} \\
    \vec{N} &= \frac{d\vec{T}}{ds} \\
    \vec{B} &= \vec{T} \times \vec{N}
\end{align*}

The change in the binormal vector is given by
\begin{align*}
    \frac{d\vec{B}}{ds} &= \frac{d(\vec{T} \times \vec{N})}{ds}
    = \frac{d\vec{T}}{ds} \times \vec{N} + \vec{T} \times \frac{d\vec{N}}{ds} \\
    &= \kappa\vec{N} \times \vec{N} + \vec{T} \times \frac{d\vec{N}}{ds} \\
    &= \vec{T} \times \frac{d\vec{N}}{ds}
\end{align*}

Clearly, \(\frac{d\vec{B}}{ds}\) is orthogonal to both \(\vec{B}\)
and \(\vec{T}\) and thus it is parallel to \(\vec{N}\).
Indeed,
\begin{align*}
    &\tau \vec{N} = -\frac{d\vec{B}}{ds} \\
    &\tau = -\frac{d\vec{B}}{ds} \cdot \vec{N}
\end{align*}
note that the minus sign is by convention.
This value is called torsion.

\section{Acceleration}

The acceleration has both a tangential and normal component.

$\begin{aligned} \vec{a} & =\frac{d \vec{v}}{d t}=\frac{d|\vec{v}|}{d t} \vec{T}+|\vec{v}| \frac{d \vec{T}}{d t} \\ & =\frac{d|\vec{v}|}{d t} \vec{T}+|\vec{v}|\left|\frac{d \vec{T}}{d t}\right| \vec{N} \\ & =\frac{d|\vec{v}|}{d t} \vec{T}+\kappa|\vec{v}|^2 \vec{N}\end{aligned}$

\section{Limits}

Limits in multivariable calculus can be approached by infinite directions.
A limit exists iff it is equal from all the directions.

\section{Differentiability}

% TODO definition

Let \(f\colon {\mathbb{R}}^m \to {\mathbb{R}}^n\) be a function.
If the partial derivatives \(\frac{\partial f}{\partial x_1}\), \(\frac{\partial f}{\partial x_2}\),
\(\cdots\), \(\frac{\partial f}{\partial x_m}\) exist and are continuous in an open region \(R\), then \(f\) is differentiable in \(R\).

\section{Multi-variable chain rule}

\pagebreak

\section{Exercises}

\subsection{Open set 1}

Prove that the set \(A=\{(x,y) \in {\mathbb{R}}^2 \,|\, 2<x^2+y^2<4\}\)
is open.

Let \(p = (x,y)\) where \(p \in A\).
The set \(A\) is open iff \(\exists \epsilon > 0 \,|\, B_\epsilon(p) \subset A \).
Let \(d = \sqrt{x^2 + y^2}\). For a radius \(\epsilon \leq \min(d-\sqrt{2}, d-\sqrt{4})\),
the open ball \(B_\epsilon(p) \subset A\).

\subsection{Norm 1}

TODO

\subsection{Countable set}

TODO % https://planetmath.org/mathbbr2setminuscispathconnectedifciscountable

\subsection{Open set 2}

Prove that the set \(A=\{(x,y) \in {\mathbb{R}}^2 \,|\, x^2<y<x\}\)
is open.

TODO

\subsection{Sequence 1}

Consider the sequence \(\{x_k\}\) in \({\mathbb{R}}^2\)
defined by \[ x_k = \left( \sin\left(\frac{\pi k}{2}\right), \frac{{(-1)}^k}{\sqrt{k}} \right) \]
for each \(k \in {\mathbb{N}}^*\).
Determine whether \(\{x_k\}\) is bounded, and if so, find a convergent subsequence
and identity its limit.

The sinusoidal part of the pair of the sequence is bounded because \(-1 \leq \sin\theta \leq 1\).
The other part has a numerator oscillating between \(1\) and \(-1\),
and the denominator goes from \(1\) to \(+\infty\) in the limit.
Thus, the sequence is absolutely decreasing and \(-1 \leq \{x_k\} \leq \frac{1}{\sqrt{2}}\).
We now notice that
\[
    \sin\left(\frac{\pi k}{2}\right)
    = \begin{cases}
        1 \text{ or } -1, & k \text{ odd} \\
        0, & k \text{ even}
    \end{cases}
\]
By considering the subsequence where \(k\) is even we get a converging sequence
\[
    \lim_{k \to \infty} \{x_{2k}\} = (0, 0) 
\]

\subsection{Level curves}

TODO: the level curve is the curve on the horizontal plane intersecting a 2 variables function

Find the equation of the level curve of the function \(f(x,y)\) that passes through
the given point p.
\begin{enumerate}
    \item \(f(x,y) = 16-x^2-y^2\) and \(p=(2\sqrt{2}, \sqrt{2})\);
    \item \(f(x,y) = \sqrt{x^2-1}\) and \(p=(1,0)\);
    \item \(f(x,y) = \int_x^y \frac{d\theta}{\sqrt{1-\theta^2}}\) and \(p=(0,1)\).
\end{enumerate}

\begin{enumerate}
    \item \(f(2\sqrt{2}, \sqrt{2}) = 6\), so the height of the plane is 6.
        By plugging \(z=6\) in we get \(6=16-x^2-y^2\)
        and thus the level curve is given by \(10=x^2+y^2\);
    \item \(f(1, 0) = 0\), so the height of the plane is 0.
        By plugging \(z=0\) in we get \(0 = \sqrt{x^2-1}\)
        and thus the level curve is given by \(x^2=1\), that is
        the lines \(x=1\) and \(x=-1\).
    \item We first solve the integral
    \[ \int_x^y \frac{d\theta}{\sqrt{1-\theta^2}} = \arcsin(y) - \arcsin(x) \]
    Now we can evaluate \(f(0,1) = \arcsin(1) - \arcsin(0) = \frac{\pi}{2}\),
    so the height of the plane is \(\frac{\pi}{2}\).
    By plugging \(z=\frac{\pi}{2}\) in we get
    \begin{align*}
        &\frac{\pi}{2} = \arcsin(y) - \arcsin(x) \\
        &y = \sin\left( \frac{\pi}{2} + \arcsin(x) \right) \\
        &y = \sqrt{1-x^2}
    \end{align*}
    with \(x < 0\).
\end{enumerate}

\subsection{Limits 1}

Solve the limit \(\lim _{(x, y) \rightarrow(2,1)} \frac{x^2-3 y}{x+2 y^2}\)

\[
    \lim _{(x, y) \rightarrow(2,1)} \frac{x^2-3 y}{x+2 y^2}
    = \frac{2^2-3}{2+2} = \frac{1}{4}
\]

\subsection{Limits 2}

Solve the limit \(\lim _{(x, y) \rightarrow(0,0)} \frac{\sin \left(x^2+y^2\right)}{x^2+y^2}\)

By using polar coordinate \(x=r\cos\theta\) and \(y=r\sin\theta\),
and thus \(x^2+y^2=r^2\), therefore
\[
    \lim_{(x, y) \rightarrow(0,0)} \frac{\sin \left(x^2+y^2\right)}{x^2+y^2}
    = \lim_{r \rightarrow 0} \frac{\sin \left(r\right)}{r}=1
\]

\subsection{Limits 3}

Solve the limit \(\lim _{(x, y) \rightarrow(0,0)} \frac{x y^2}{x^2+y^4}\)

\subsection{Limits 4}

Solve the limit \(\lim _{(x, y) \rightarrow(0,0)} \frac{x^2 y}{x^2+y^4}\)

\end{document}