\documentclass[a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{blkarray}
\usepackage{tikz}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={Graphs},
    pdfpagemode=FullScreen,
}

\title{Graphs}
\author{Paolo Bettelini}
\date{}

\begin{document}

\maketitle
\tableofcontents
\pagebreak

\section{Definition}

A \textit{graph} \(G=(V,E)\) is a pair consisting of a set V (vertices)
and a set \(E\) (edges). Every element in \(E\) is a distinct pair of vertices
in \(V\).

\section{Degree}

The \textit{degree} of a vertex \(v\), \(deg(v)\) is defined
as the numbers of edges that are incident on \(v\).

The sum of the degrees of all vertices is equals to twice the number of edges.
\[
    \sum_{v\in V}deg(v)=2|E|
\]

\section{Paths}

A \textit{path} is a sequence of vertices connected by edges.
\\A path is a \(cycle\) if it starts and ends at the same vertex.
\\A path is \textit{simple} if every vertex in the path is distinct.
\\A graph is \textit{connected} if there is at least a path
between any pair of vertices.

\section{Adjacency Matrices}

A finite graph can be represented by a square matrix
\(n \times n\) where \(n\) is the number of vertices.

\begin{center}
    \begin{minipage}[r]{6cm}
        \begin{tikzpicture}[main/.style = {draw, circle}]
            \foreach \x [count = \xi] in {a,b,c,d,e} {
                \node[main] (\xi) at
                    ({1 * cos(360 / 5 * \xi)}, {1 * sin(360 / 5 * \xi)})
                    {\(\x\)};
            }
        
            \draw (1) -- (2);
            \draw (1) -- (3);
            \draw (1) -- (5);
            \draw (2) -- (4);
            \draw (5) -- (3);
        \end{tikzpicture} 
    \end{minipage}
    \begin{minipage}[l]{6cm}
        A=
        \begin{blockarray}{cccccc}
            & a & b & c & d & e \\
            \begin{block}{c(ccccc)}
            a & 0 & 1 & 1 & 0 & 1 \\
            b & 1 & 0 & 0 & 1 & 0 \\
            c & 1 & 0 & 0 & 0 & 1 \\
            d & 0 & 1 & 0 & 0 & 0 \\
            e & 1 & 0 & 1 & 0 & 0 \\
            \end{block}
        \end{blockarray}
    \end{minipage}
\end{center}

Every row and column represents a vertice. \(1\) means
that the two vertices are adjacent, \(0\) otherwise. The diagonal of this matrix
will always e \(0s\) since no vertice is adjacent to itself
and \(A=A^t\)

\end{document}
